Algebraic File Synchronization: Adequacy and Completeness

### 2016-10-08 p2/works-inverse.tex
New symbol and definition for the w relation

\begin{mydef}[$\works{x}$]
% Unsure if this relation should be in the algebra.
For a set of $k$ sequences
$\works{A_1,A_2,\ldots,A_k}$ means that 
$A_1,A_2,\ldots,A_k$ work at the same time, that is,
% intended interpretation:
\[\exists \FS: A_1\FS\neq\fsbroken \wedge \cdots \wedge A_k\FS\neq\fsbroken.\]
As the sequences form a set, their order is irrelevant.
% inference rules:
We also know that $\works{A} \Leftrightarrow A\nequiv \cbrk$. 
\end{mydef}

\begin{mydef}[$\worksc{x}{y}$]
For two sets of sequences, $\worksc{A_1,A_2,\ldots,A_k}{B_1,B_2,\ldots,B_l}$ means that 
all of $A_1,\ldots,A_k$ work where all of $B_1,\ldots,B_l$ work,
that is,
% intended interpretation:
\begin{align*}
\forall \FS:{}& 
B_1\FS\neq\fsbroken \wedge \cdots \wedge B_k\FS\neq\fsbroken\\
&\Rightarrow\\
&A_1\FS\neq\fsbroken \wedge \cdots \wedge A_l\FS\neq\fsbroken.
\end{align*}
As above, the order of the sequences in the sets is not relevant.
\end{mydef}

@@

@@

% An inference rule in the algebra
% $\worksc{X}{X}$ is always true for any set of sequences $X$,

% \begin{mycor}
% % An inference rule in the algebra
% $B\eqext A \Rightarrow \worksc{A}{B}$.
% \end{mycor}

\begin{mycor}\label{worksextpostfix}
% An inference rule in the algebra
$\forall A,S: \worksc{A}{A\cc S}$, that is, if a sequence works, its initial segment also works.
\end{mycor}

\begin{mycor}\label{workschained}
% An inference rule in the algebra
$\workssign$ can be chained:
if $\seqset{A'}\subset\seqset{A}$, then
$ \worksc{\seqset{C}}{\seqset{A'}} \wedge \worksc{\seqset{A}}{\seqset{B}} \Rightarrow \worksc{\seqset{A}\cup\seqset{C}}{\seqset{B}}$.
\end{mycor}

@@

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\[ \alpha\indep \beta \Rightarrow \worksc{\alpha\cc \beta}{\alpha, \beta}. \]
% \begin{align*}
% \forall\FS:{}&\cxynv\indep \czwmv \\
% &\quad\wedge \cxynv\aFS\neq\fsbroken \\
% &\quad\wedge \czwmv\aFS\neq\fsbroken \\
% &\Rightarrow \cxynv\cc \czwmv\aFS\neq\fsbroken.
% \end{align*}
\end{myax}
\begin{proof}
We name this proposition a \namecref{combine_independent_commands} 
because to prove it, we must reach back to our filesystem model.
Let $\alpha=\cxynv$ and $\beta=\czwmv$.
%% TODO by contradiction
We proceed in an indirect way and
assume that there is a filesystem $\FS$ for which
$\cxynv\aFS\neq\fsbroken$ and $\czwmv\aFS\neq\fsbroken$, but
$\cxynv\cc \czwmv\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $[\FS(m)]=[w]$ and so
$[(\cxynv\aFS)(m)]=[w]$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

@@

\begin{mylem}\label{combine_independent_sequences}
The combination of independent sequences works wherever the original sequences work:
\[ S\indep T \Rightarrow \worksc{S\cc T}{S,T}. \]
\end{mylem}
\begin{proof}
Assume that there is a filesystem $\FS$ so that
$S\aFS\neq\fsbroken$ and $T\aFS\neq\fsbroken$, but
$(S\cc T)\FS=\fsbroken$.

@@

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\works{A,B}$,
and there are commands on node $n$ in both $A$ ($\cxynv\in A$) and $B$ ($\czwnv\in B$),
then the input types of these commands must match ($[x]=[z]$).
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands} and
%% TODO by contradiction
is easily shown using an indirect proof: if $[x]\neq [z]$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $A$ and $B$ cannot work on the same filesystem.
\end{proof}

@@

@@

\begin{mylem}\label{r_invmove}
A common initial segment of a set of \emph{representative} sequences can be moved to the other side of $\workssign$ by inverting it:
\begin{gather*}
\worksc{B\cc \seqset{A}}{\seqset{C}} \Rightarrow \worksc{\seqset{A}}{B^{-1}\cc \seqset{C}} \\
\worksc{B}{\seqset{C}} \wedge \worksc{\seqset{A}}{B^{-1}\cc \seqset{C}} \Rightarrow \worksc{B\cc \seqset{A}}{\seqset{C}}
\end{gather*}
\end{mylem}
\begin{proof}

The proof for both propositions is illustrated by \cref{fig_invmove}, where
$\Dom{X}$, the domain of $X$, represents the set of filesystems
the sequence $X$ does not break, or neither of the sequences in the set $X$ breaks.
We write $\Img{X}$ to denote the image of the domain of the sequence $X$.
We also note that in general $\worksc{\seqset{X}}{\seqset{Y}} \Leftrightarrow \Dom{\seqset{Y}} \subset \Dom{\seqset{X}}$.

@@

Unless $\works{\seqset{C}}$, both propositions are trivially true. Otherwise,
for the first proposition
we know $\works{B\cc\seqset{A}}$, which implies that $\Dom{\seqset{A}}$ intersects with $\Img{B}$ 
as otherwise $B\cc\seqset{A}$ would break all filesystems.
As the (representative) $B$ is a bijection between its domain and image,
we can use $B^{-1}$ to map this intersection back
onto the domain of $B$ and get $\Dom{B\cc\seqset{A}}$.
As $\worksc{B\cc \seqset{A}}{\seqset{C}}$, 
we know that $\Dom{\seqset{C}} \subset \Dom{B\cc\seqset{A}}$.
We now use $B$ to map both sides of this statement, which yields
$\Dom{B^{-1}\cc \seqset{C}} \subset \Dom{B^{-1}\cc B\cc\seqset{A}}$,
but the latter is a subset of $\Dom{\seqset{A}}$,
and so $\Dom{B^{-1}\cc \seqset{C}} \subset \Dom{\seqset{A}}$,
from which $\worksc{\seqset{A}}{B^{-1}\cc \seqset{C}}$.

For the second proposition, 
from $\worksc{B}{\seqset{C}}$ we know $\Dom{\seqset{C}} \subset \Dom{B}$.
Therefore we can project the whole of $\Dom{\seqset{C}}$ using the bijective representative sequence $B$
to get $\Dom{B^{-1}\cc\seqset{C}}$, and as $\worksc{\seqset{A}}{B^{-1}\cc \seqset{C}}$
we know
$\Dom{B^{-1}\cc\seqset{C}} \subset \Dom{\seqset{A}}$.
We use $B^{-1}$ to project both sides of this statement to get
$\Dom{B\cc B^{-1}\cc\seqset{C}} \subset \Dom{B\cc\seqset{A}}$.
But we also know that $\Dom{B\cc B^{-1}\cc\seqset{C}} = \Dom{\seqset{C}}$
as $\Dom{\seqset{C}} \subset \Dom{B}$,
and so $\Dom{\seqset{C}} \subset \Dom{B\cc\seqset{A}}$
from which $\worksc{B\cc \seqset{A}}{\seqset{C}}$.
\end{proof}

>>>>

@@

\begin{mydef}[$\Dom{X}$ and $\Img{X}$]
$\Dom{X}$ is the domain of $X$.
For a sequence of commands $A$, $\Dom{A}$ is the set of filesystems $A$ does not break.
For a set of sequences $\seqset{A}$, $\Dom{\seqset{A}}$ is the set of filesystems no $A\in\seqset{A}$ breaks.
We write $\Img{X}$ to denote the image of the domain of the sequence $X$.
\end{mydef}

\begin{mydef}[$\worksmsign$, $\worksmeqsign$]
For two sets of sequences $\sqs{A}$ and $\sqs{B}$
we write $\worksm{\sqs{A}}{\sqs{B}}$ iff $\Dom{\sqs{A}} \subseteq \Dom{\sqs{B}}$;
that is, if all sequences in $\sqs{B}$ work (do not break the filesystem)
on any filesystem all sequences in $\sqs{A}$ work on.
We write $\worksmstr{\sqs{A}}{\sqs{B}}$ iff $\Dom{\sqs{A}} \subset \Dom{\sqs{B}}$.

As a special case, we write $\worksmnil{\sqs{A}}$ if we know that there is a filesystem
all sequences in $\sqs{A}$ work on,
and we apply these relations to sequences directly:
$\worksm{A}{B}$ with the meaning $\worksm{\{A\}}{\{B\}}$.
\end{mydef}

@@

\begin{mycor}\label{worksextpostfix}
% An inference rule in the algebra
$\forall A,S: \worksm{A\cc S}{A}$, that is, if a sequence works, its initial segment also works.
\end{mycor}

\begin{mycor}\label{workschained}
% An inference rule in the algebra
$\worksmeqsign$ can be chained:
if $\seqset{B'}\subseteq\seqset{B}$, then
$\worksm{\sqs{A}}{\sqs{B}} \wedge \worksm{\sqs{B'}}{\sqs{C}} \Rightarrow \worksm{\sqs{A}}{\sqs{B}\cup\sqs{C}}$.
\end{mycor}

@@

\begin{myax}\label{combine_independent_commands}
The combination of independent commands works wherever the original commands work:
\[ \alpha\indep \beta \Rightarrow \worksm{\{\alpha, \beta\}}{\{\alpha\cc \beta\}}. \]
\end{myax}
\begin{proof}
We name this proposition a \namecref{combine_independent_commands} 
because to prove it, we must reach back to our filesystem model.
Let $\alpha=\cxynv$ and $\beta=\czwmv$.
%% TODO by contradiction
We proceed in an indirect way and
assume that there is a filesystem $\FS$ for which
$\cxynv\aFS\neq\fsbroken$ and $\czwmv\aFS\neq\fsbroken$, but
$\cxynv\cc \czwmv\aFS=\fsbroken$.
We know $\cxynv\aFS\neq\fsbroken$ so it must be applying 
$\czwmv$ that breaks it.
Applying a command can only result in a broken filesystem in three cases.
First, if the filesystem was already broken, which cannot be the case here.
Second, if the input type does not match the filesystem,
but we know $[\FS(m)]=[w]$ and so
$[(\cxynv\aFS)(m)]=[w]$ as based on \cref{incomparable_is_independent}, $n\neq m$.
Third, if the new filesystem violates the tree property.
This again cannot be the case because we also know that $n\unrel m$
and the tree property only depends on the types of the parent and children of $m$,
which therefore cannot be changed by $\cxynv$.
\end{proof}

@@

\begin{mylem}\label{combine_independent_sequences}
The combination of independent sequences works wherever the original sequences work:
\[ S\indep T \Rightarrow \worksm{\{S,T\}}{\{S\cc T\}}. \]
\end{mylem}
\begin{proof}
Assume that there is a filesystem $\FS$ so that
$S\aFS\neq\fsbroken$ and $T\aFS\neq\fsbroken$, but
$(S\cc T)\FS=\fsbroken$.

@@

\begin{mylem}\label{worksinputmatch}
If $A$ and $B$ are minimal sequences, $\worksmnil{\{A,B\}}$,
and there are commands on node $n$ in both $A$ ($\cxynv\in A$) and $B$ ($\czwnv\in B$),
then the input types of these commands must match ($[x]=[z]$).
\end{mylem}
\begin{proof}
This result is similar to \cref{equiv_simple_same_commands} and
%% TODO by contradiction
is easily shown using an indirect proof: if $[x]\neq [z]$, then there is no filesystem that
either $\cxynv$ or $\czwnv$ would not break, 
and consequently $A$ and $B$ cannot work on the same filesystem.
\end{proof}

@@

% TODO We introduced BREAK to deal with cases where commands
% are in effect not defined. Now we need to undo this 
% to see that repr commands are bijective.

@@

\begin{mylem}\label{r_invmove}
A common initial segment of a set of \emph{representative} sequences can be moved to the other side of $\worksmeqsign$ by inverting it:
\begin{gather*}
\worksm{\seqset{C}}{B\cc \seqset{A}}
\ \Longleftrightarrow \ \worksm{B^{-1}\cc \seqset{C}}{\seqset{A}} 
\ \wedge \  \worksm{\seqset{C}}{\{B\}}
\end{gather*}
\end{mylem}
\begin{proof}
% TODO Rewrite for equivalence, and swap A and C.
% This result is quite expected as we're mapping a subset statement using a bijection,
% but made more complicated a as B is not defined everywhere.

The proof for both propositions is illustrated by \cref{fig_invmove}.

@@

Unless $\works{\seqset{C}}$, both propositions are trivially true. Otherwise,
for the first proposition
we know $\works{B\cc\seqset{A}}$, which implies that $\Dom{\seqset{A}}$ intersects with $\Img{B}$ 
as otherwise $B\cc\seqset{A}$ would break all filesystems.
As the (representative) $B$ is a bijection between its domain and image,
we can use $B^{-1}$ to map this intersection back
onto the domain of $B$ and get $\Dom{B\cc\seqset{A}}$.
As $\worksc{B\cc \seqset{A}}{\seqset{C}}$, 
we know that $\Dom{\seqset{C}} \subseteq \Dom{B\cc\seqset{A}}$.
We now use $B$ to map both sides of this statement, which yields
$\Dom{B^{-1}\cc \seqset{C}} \subseteq \Dom{B^{-1}\cc B\cc\seqset{A}}$,
but the latter is a subseteq of $\Dom{\seqset{A}}$,
and so $\Dom{B^{-1}\cc \seqset{C}} \subseteq \Dom{\seqset{A}}$,
from which $\worksc{\seqset{A}}{B^{-1}\cc \seqset{C}}$.

For the second proposition, 
from $\worksc{B}{\seqset{C}}$ we know $\Dom{\seqset{C}} \subseteq \Dom{B}$.
Therefore we can project the whole of $\Dom{\seqset{C}}$ using the bijective representative sequence $B$
to get $\Dom{B^{-1}\cc\seqset{C}}$, and as $\worksc{\seqset{A}}{B^{-1}\cc \seqset{C}}$
we know
$\Dom{B^{-1}\cc\seqset{C}} \subseteq \Dom{\seqset{A}}$.
We use $B^{-1}$ to project both sides of this statement to get
$\Dom{B\cc B^{-1}\cc\seqset{C}} \subseteq \Dom{B\cc\seqset{A}}$.
But we also know that $\Dom{B\cc B^{-1}\cc\seqset{C}} = \Dom{\seqset{C}}$
as $\Dom{\seqset{C}} \subseteq \Dom{B}$,
and so $\Dom{\seqset{C}} \subseteq \Dom{B\cc\seqset{A}}$
from which $\worksc{B\cc \seqset{A}}{\seqset{C}}$.
\end{proof}
